---
title: "03-validate-models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Libraries and Data

```{r}
library(brms)
library(rstan)
library(tidyverse)
library(cowplot)
library(colorblindr)
library(ggpubr)
library(tidybayes)
library(modelr)

# recommended code to speed up stan
# see https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
vis_display_order <- c("point","barchart","interval","table","hopsdist","dotplot","hops","density")
max_eval = 30

rq3 <- read_csv("data/rq3.csv") %>%
    mutate(
    eval_period_normalized = ((evalPeriod - max_eval) / max_eval) + 0.5
  )
```

```{r fig.width=10,fig.height=4}
rq3 %>%
  ggplot(aes(x = evalPeriod, y = dv, color = treatment)) +
  stat_summary(fun.data = mean_se) +
  geom_hline(yintercept = 1) +
  stat_smooth(method = lm) +
  facet_grid(.~treatment) +
  theme(legend.position = "none")
```

### Priors

Commented out.

```{r}
get_prior(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment
  ),
  data = rq3, family = Beta)
```


```{r}
pr_beta = c(
  prior(normal(0, 1), class = b),
  # these prior intercepts are wide and cover 0 (50% on the logit scale), but
  # also assume some likely better-than-50% performance on average --- this
  # was chosen to aid convergence during the pilot, but does not have a strong
  # impact on final estimates.
  prior(normal(2, 2), class = Intercept),
  prior(normal(2, 2), class = Intercept, dpar = phi),
  prior(normal(0, 1), class = b, dpar = phi),
  prior(student_t(3, 0, 1), class = sd)
)
```

### Fit model

```{r run, eval=FALSE, include=TRUE}
# save model - core model
   # dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
   #  phi ~ eval_period_normalized * treatment),
mbeta = brm(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment),
  data = rq3, 
  prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )


save(mbeta, file = "models/fit1.rda")
```

```{r run2, eval=FALSE, include=TRUE}
mbeta2 = brm(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment),
  data = rq3, 
  prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )   #  phi ~ eval_period_normalized ),
# challenger: phi eval_period_nor
save(mbeta2, file = "models/fit2.rda")
```


```{r run3, eval=FALSE, include=TRUE}
mbeta3 = brm(bf(
    dv ~ eval_period_normalized + treatment + (1|usertoken), 
    phi ~ eval_period_normalized),
  data = rq3, 
  prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )   #  phi ~ eval_period_normalized ),
# challenger: phi eval_period_nor
save(mbeta3, file = "models/fit3.rda")
```



```{r run4, eval=FALSE, include=TRUE}
# save model - core model
   # dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
   #  phi ~ eval_period_normalized * treatment),
mbeta4 = brm(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment),
  data = rq3, 
  #prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )


save(mbeta4, file = "models/fit4.rda")
```

```{r run5, eval=FALSE, include=TRUE}
mbeta5 = brm(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment),
  data = rq3, 
  #prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )   #  phi ~ eval_period_normalized ),
# challenger: phi eval_period_nor
save(mbeta5, file = "models/fit5.rda")
```


```{r run6, eval=FALSE, include=TRUE}
mbeta6 = brm(bf(
    dv ~ eval_period_normalized + treatment + (1|usertoken), 
    phi ~ eval_period_normalized),
  data = rq3, 
  #prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )   #  phi ~ eval_period_normalized ),
# challenger: phi eval_period_nor
save(mbeta6, file = "models/fit6.rda")
```




```{r}
# load model
load("models/fit1.rda")
load("models/fit2.rda")
load("models/fit3.rda")
load("models/fit4.rda")
load("models/fit5.rda")
load("models/fit6.rda")
```


## Model results

```{r}
summary(mbeta)
```

```{r}
summary(mbeta2)
```

```{r}
summary(mbeta3)
```

```{r}
summary(mbeta4)
```

```{r}
summary(mbeta5)
```

```{r}
summary(mbeta6)
```

## PPC across models

```{r}
pp_check(mbeta)
```

```{r}
pp_check(mbeta2)
```

```{r}
pp_check(mbeta3)
```

```{r}
pp_check(mbeta4)
```

```{r}
pp_check(mbeta5)
```
```{r}
pp_check(mbeta6)
```

## Conditional Effects

```{r}
conditional_effects(mbeta, "treatment") # conditions = data.frame(size = 1)
```


```{r}
conditional_effects(mbeta2, "treatment") # conditions = data.frame(size = 1)
```

```{r}
conditional_effects(mbeta3, "treatment") # conditions = data.frame(size = 1)
```

```{r}
conditional_effects(mbeta4, "treatment") # conditions = data.frame(size = 1)
```

```{r}
conditional_effects(mbeta5, "treatment") # conditions = data.frame(size = 1)
```

```{r}
conditional_effects(mbeta6, "treatment") # conditions = data.frame(size = 1)
```



## Mean PPC

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta2, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta3, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta4, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta5, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```

```{r}
#q75 <- function(y) quantile(y, 0.75)
pp_check(mbeta6, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```


```{r}
bayesplot::pp_check(mbeta, type = "violin_grouped", group = "treatment")
```


```{r}
bayesplot::pp_check(mbeta, type = "error_hist")
```

## Plots


```{r}
eval_period_ends = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```




```{r}
eval_period_ends2 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta2, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends2 %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```


Notice how when we remove interaction density seems to be significant. This is because we're incorrectly applying the population's evaluation curve.


```{r}
eval_period_ends3 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta3, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends3 %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```

```{r}
eval_period_ends4 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta4, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends4 %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```



```{r}
eval_period_ends5 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta5, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends5 %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```

Again, without interaction we're having issues. 

```{r}
eval_period_ends6 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta6, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))


#Conditional means (for "average" person on "average" scenario) on last trial:
eval_period_ends6 %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 1.3) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
```

```{r}
max_eval=30

mu_mean = rq3 %>%
  mutate(
    eval_period_normalized = ((evalPeriod - max_eval) / max_eval) + 0.5
  ) %>%
  group_by(treatment, eval_period_normalized) %>%
  summarise(mean=mean(dv))
```


```{r}
fit_lines6 = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = seq_range(eval_period_normalized, n = 20)
  ) %>%
  add_fitted_samples(mbeta6, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup()  %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))

mu_mean6 = fit_lines6 %>%
  group_by(treatment, eval_period_normalized) %>%
  summarise(mu_mean=mean(mu))
```

```{r}
fit_lines = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = seq_range(eval_period_normalized, n = 20)
  ) %>%
  add_fitted_samples(mbeta, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() 
  #mutate(vis = fct_relevel(vis, vis_display_order))
```


As we see here, the model is not doing well. We need the interaction effect.

```{r, fig.height = 3, fig.width = 8}
scale_fill_fit_lines = scale_fill_manual(
  values = RColorBrewer::brewer.pal(4, "Greys")[-1], guide = guide_legend(reverse = TRUE)
)
fit_lines %>%
  ggplot(aes(x = eval_period_normalized, y = mu)) +
  stat_lineribbon(.prob = c(.95, .8, .5)) +
  geom_hline(yintercept = seq(.7, 1, by=.05), color="gray75", alpha = 0.5) +
  geom_line(aes(y = mean), data = mu_mean, color = "red", size = 1) +
  facet_grid(. ~ forcats::fct_relevel(treatment, vis_display_order)) +
  scale_fill_fit_lines 
  #fit_line_plot_settings
```

# This is the actual model

```{r, fig.height = 3, fig.width = 8}
scale_fill_fit_lines = scale_fill_manual(
  values = RColorBrewer::brewer.pal(4, "Greys")[-1], guide = guide_legend(reverse = TRUE)
)
fit_lines %>%
  ggplot(aes(x = eval_period_normalized, y = mu)) +
  stat_lineribbon(.prob = c(.95, .8, .5)) +
  geom_hline(yintercept = seq(.7, 1, by=.05), color="gray75", alpha = 0.5) +
  geom_line(aes(y = mean), data = mu_mean, color = "red", size = 1) +
  facet_grid(. ~ forcats::fct_relevel(treatment, vis_display_order)) +
  scale_fill_fit_lines 
  #fit_line_plot_settings
```