---
title: "04-analyze-modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## R Markdown

```{r}
library(brms)
library(rstan)
library(tidyverse)
library(cowplot)
library(colorblindr)
library(ggpubr)
library(tidybayes)
library(modelr)

# recommended code to speed up stan
# see https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

max_eval = 30

rq3 <- read_csv("data/rq3.csv") %>%
    mutate(
    eval_period_normalized = ((evalPeriod - max_eval) / max_eval) + 0.5
  )
```

```{r fig.width=10,fig.height=4}
rq3 %>%
  ggplot(aes(x = evalPeriod, y = dv, color = treatment)) +
  stat_summary(fun.data = mean_se) +
  geom_hline(yintercept = 1) +
  stat_smooth(method = lm) +
  facet_grid(.~treatment) +
  theme(legend.position = "none")
```

### Priors

Commented out.

```{r}
get_prior(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment
  ),
  data = rq3, family = Beta)
```


```{r}
pr_beta = c(
  prior(normal(0, 1), class = b),
  # these prior intercepts are wide and cover 0 (50% on the logit scale), but
  # also assume some likely better-than-50% performance on average --- this
  # was chosen to aid convergence during the pilot, but does not have a strong
  # impact on final estimates.
  prior(normal(2, 2), class = Intercept),
  prior(normal(2, 2), class = Intercept, dpar = phi),
  prior(normal(0, 1), class = b, dpar = phi),
  prior(student_t(3, 0, 1), class = sd)
)
```

### Fit model

Let's fit the model:

```{r, eval = FALSE}
warmup = 2000
iter = warmup + 2000
thin = 2
mbeta = brm(bf(
    dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
    phi ~ eval_period_normalized * treatment),
  data = rq3, 
  prior = pr_beta, 
  control = list(adapt_delta = 0.9995, max_treedepth = 15, stepsize = 0.005),
  warmup = warmup, iter = iter, thin = thin,
  family = Beta
  )

# message https://discourse.mc-stan.org/t/dealing-with-catalina/11285/254
```


```{r}
# pairs(mbeta$fit, pars = c("b_Intercept", "b_trial_normalized", "sd_participant__Intercept",  "sd_participant__trial_normalized", 
#   "sd_scenario__Intercept",  
#   "cor_participant__Intercept__trial_normalized", "b_phi_Intercept", "b_phi_trial_normalized"))
```

```{r run, eval=FALSE, include=TRUE}
# save model - core model
   # dv ~ eval_period_normalized + treatment + eval_period_normalized * treatment + (1|usertoken), 
   #  phi ~ eval_period_normalized * treatment),
save(mbeta, file = "models/fit1.rda")
```



```{r}
# load model
load("models/fit1.rda")
```

```{r}
summary(mbeta)
```

```{r}
treatment_gg <- conditional_effects(mbeta, "treatment") # conditions = data.frame(size = 1)

treatment_gg
```


```{r}

conditional_effects(mbeta) # conditions = data.frame(size = 1)

```


```{r}
plot(mbeta)
```

```{r}
#get_variables(mbeta)
```

```{r}
pp_check(mbeta)
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```


```{r}
#q75 <- function(y) quantile(y, 0.75)
#pp_check(mbeta, type='stat', stat='mean')
#+scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30"))
```



```{r}
bayesplot::pp_check(mbeta, type = "violin_grouped", group = "treatment")
```


```{r}
bayesplot::pp_check(mbeta, type = "error_hist")
```


## Results

```{r}
library(tidybayes)
library(modelr)

pred_beta = 
  rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = seq_range(eval_period_normalized, n = 20)
  ) %>%
  add_predicted_samples(mbeta, re_formula = NULL, allow_new_levels = TRUE) %>%    
  mean_qi(.prob = c(.95, .8, .5))
```


```{r fig.height = 2, fig.width = 6}
# shared properties of posterior prediction and fit line plots
fit_line_plot_settings = list(
  scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30")),
  coord_cartesian(expand = FALSE),
  xlab("Evaluation Period"),
  theme(
    panel.grid = element_blank(), panel.spacing.x = unit(5, "points"),
    strip.background = element_blank(), strip.text = element_text(hjust = 0.5, color = "black"),
    axis.title.x = element_text(hjust = 0)
  ))



vis_display_order <- c("point","barchart","interval","table","hopsdist","dotplot","hops","density")
labs_vis = c("Point + Interval","Bar Chart","Interval","Table","HOP + Strip","Dot plot","HOP","Probability Density")
labeller = list(labs_vis)

vis_labels = list(
  "point"="Point + Interval",
  "barchart"="Bar Chart",
  "interval"="Interval",
  "table"="Table",
  "hopsdist"="HOP + Strip",
  "dotplot"="Dot plot",
  "hops"="HOP",
  "density"="Prob. Density"
)

vis_labeller <- function(variable,value){
  return(vis_labels[value])
}
#vnames(labs_vis) = vis_display_order

post_pred_plot = pred_beta %>%
  ungroup() %>%
  #mutate(treatment = forcats::fct_relevel(treatment, vis_display_order)) %>%
  ggplot(aes(x = eval_period_normalized)) +
  geom_lineribbon(aes(y = pred), data = pred_beta) +
  geom_line(aes(y = pred), data = pred_beta, color = "red",size = 1) +
  geom_hline(yintercept = 1) +
  scale_fill_brewer(guide = guide_legend(reverse = TRUE)) +
  geom_hline(yintercept = seq(.3, .95, by=.1), color="gray75", alpha = 0.5) +
  fit_line_plot_settings + 
  #facet_grid(. ~ treatment, labeller = labeller(treatment = labs_vis)) +
  facet_grid(. ~ forcats::fct_relevel(treatment, vis_display_order), labeller = vis_labeller) +
  labs(title = " ", y = "") +
  geom_point(aes(y = dv), alpha = 0.05, data = rq3) 

post_pred_plot
```

### code below needs to be modified

Above, the red line is the predicted median, and the blue bands are predictive intervals for the data. We can see that most conditions have a distance between 1 year and 30 year evaluation periods which indicate myopic loss aversion. However, there are some differences, and it is hard to tell how reliable those differences are just by looking at predictions. So, let's look at posteriors for the mean and precision of estimates according to the model.


## Model-based myopic loss aversion

First we'll generate samples of fit lines for conditional means and variances. We will use these to plot fit lines and to generate estimates for performance in the 1 and 30 year evaluation periods. These estimates will be for the "average" scenario and "average" person:


```{r}
fit_lines = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = seq_range(eval_period_normalized, n = 20)
  ) %>%
  add_fitted_samples(mbeta, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() 
  #mutate(vis = fct_relevel(vis, vis_display_order))
```

The estimates of $\phi$ (`phi`, the precision parameter of the beta distribution) are a little hard to interpret, so instead we'll derive a posterior distribution for standard deviation. We can use the fact that the standard deviation $\sigma$ of a Beta distribution is:

$$
\sigma = \sqrt{\frac{\mu (1 - \mu)}{(1 + \phi)}}
$$

Thus we can transform samples from the distribution of $\mu$ (`mu`) and $\phi$ (`phi`) into samples from the distribution of $\sigma$ (`sd`):

```{r}
fit_lines <- fit_lines %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))
```


### Trends in mean

```{r}
mu_mean = fit_lines %>%
  group_by(treatment, eval_period_normalized) %>%
  summarise(mu_mean=mean(mu))
```


Estimates of the mean for the "average" scenario and "average" person:

```{r, fig.height = 3, fig.width = 8}
scale_fill_fit_lines = scale_fill_manual(
  values = RColorBrewer::brewer.pal(4, "Greys")[-1], guide = guide_legend(reverse = TRUE)
)
mu_lines_plot = fit_lines %>%
  ggplot(aes(x = eval_period_normalized, y = mu)) +
  stat_lineribbon(.prob = c(.95, .8, .5)) +
  geom_hline(yintercept = seq(.7, 1, by=.05), color="gray75", alpha = 0.5) +
  geom_line(aes(y = mu_mean), data = mu_mean, color = "red", size = 1) +
  facet_grid(. ~ forcats::fct_relevel(treatment, vis_display_order), labeller = vis_labeller) +
  scale_fill_fit_lines +
  fit_line_plot_settings
mu_lines_plot
```

### Trends in standard deviation

```{r}
sd_mean = fit_lines %>%
  group_by(treatment, eval_period_normalized) %>%
  summarise(sd_mean=mean(sd))
```

Estimates of the standard deviation for the "average" scenario and "average" person:

```{r, fig.height = 3, fig.width = 8}
fit_line_plot_2 = list(
  scale_x_continuous(breaks = seq(-0.5, 0.5, length.out = 7), labels = c("1", "5","10","15","20","25", "30")),
  coord_cartesian(expand = FALSE),
  xlab("Evaluation Period"),
  theme(
    panel.grid = element_blank(), 
    panel.spacing.x = unit(5, "points"),
    strip.background = element_blank(), 
    strip.text = element_text(hjust = 0.5, color = "black"),
    axis.title.x = element_text(hjust = 0)
  ))


sd_lines_plot = fit_lines %>%
  ggplot(aes(x = eval_period_normalized, y = sd)) +
  stat_lineribbon(.prob = c(.95, .8, .5)) +
  geom_hline(yintercept = seq(0, .2, by=.05), color="gray75", alpha = 0.5) +
  geom_line(aes(y = sd_mean), data = sd_mean, color = "red", size = 1) +
  facet_grid(. ~ forcats::fct_relevel(treatment, vis_display_order), labeller = vis_labeller) +
  scale_fill_fit_lines +
  fit_line_plot_2
sd_lines_plot
```



## Convert DV to Dollar for hypothetical situation

```{r}
dv_returns <- read_csv("data/dv-returns.csv")
```

```{r}
optimal_expreturn <- max(dv_returns$ExpReturn)
```

```{r}
getExpectedInvestment <- function(dv, investment, optimal_expreturn){
  return = dv * optimal_expreturn
  scales::dollar(investment * ((1 + return) ^ 30))
}

# 35 year old with $50,000 investment for different dv's
dollarValue = getExpectedInvestment(seq(0.4,1,0.2),50000, optimal_expreturn)
```

## Final figure

```{r, fig.width = 8, fig.height = 8, include = FALSE}
# generate combined plot as a PDF (this is for the paper, don't include in md output)
#dollarValue1 = getExpectedInvestment(seq(0.4,1,0.2),50000)

lines_plots = plot_grid(
  post_pred_plot +
    xlab(NULL) +
    ylab("\n\n") +
    cowplot::theme_minimal_hgrid() +
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = seq(.3, 1.0, by = .1), labels = scales::number_format(accuracy = 0.01)) + 
    #  labels = formatC(seq(.3, 1.0, by = .1), format = "f", digits = 2)) +
    #scale_y_continuous(breaks = seq(0.4,1,0.2), labels = getExpectedInvestment(seq(0.4,1,0.2),50000, optimal_expreturn)) +
    theme(
      panel.border = element_rect(color = "gray75", fill = NA),
       plot.background = element_rect(fill = 'white', color = 'white'),
      axis.text.x = element_text(size=8),
      axis.text.y = element_text(size=10),
      strip.text.x = element_text(size = 8),
      axis.title.y = element_text(size=12)
    ) ,
  mu_lines_plot +
    xlab(NULL)+
        cowplot::theme_minimal_hgrid() +
    ylab("\n\n")+
    theme(
      #strip.background = element_blank(), 
      #strip.text = element_blank(),
      strip.text.x = element_text(size = 8),
       plot.background = element_rect(fill = 'white', color = 'white'),
      axis.text.y = element_text(size=10),
      axis.text.x = element_text(size=8),
      axis.title.y = element_text(size=12)
    ) +
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = seq(.7, 1.0, by = .1), labels = scales::number_format(accuracy = 0.01)) +
    #  labels = formatC(seq(.7, 1.0, by = .1), format = "f", digits = 2)) +
    #scale_y_continuous(breaks = seq(0.7,1,0.1), labels = getExpectedInvestment(seq(0.7,1,0.1),50000, optimal_expreturn))+
    theme(panel.border = element_rect(color = "gray75", fill = NA)),
  sd_lines_plot +
    ylab("\n\n")+
        cowplot::theme_minimal_hgrid() +
    #scale_y_continuous(breaks = seq(0,.15,0.05))+
    theme(#strip.background = element_blank(), 
      #strip.text = element_blank(),
      strip.text.x = element_text(size = 8)) +
    theme(
      panel.border = element_rect(color = "gray75", fill = NA),
       plot.background = element_rect(fill = 'white', color = 'white'),
      axis.text.y = element_text(size=10),
      axis.text.x = element_text(size=8),
      axis.title.y = element_text(size=12)
       #plot.margin = margin(0, 0, 0, 0, "cm"),
      ),
  align = "h",
  ncol = 1,
  axis = "l"
)


lines_plots
pdf(file = "img/rq/lines_plots.pdf", useDingbats = FALSE, width = 10, height = 8)
lines_plots
dev.off()
```

### compare 1 vs 30


## Performance on last trial

1 year and 30 year evaluation periods.

```{r}
eval_period_ends = rq3 %>%
  data_grid(
    treatment,
    eval_period_normalized = c(-0.467, .5) #c(-0.467, .5)  # because we normalized trial to be from -0.5 to 0.5
  ) %>%
  add_fitted_samples(mbeta, re_formula = NA, var = "mu", dpar = TRUE) %>%
  ungroup() %>%
  mutate(treatment = fct_rev(fct_relevel(treatment, vis_display_order))) %>%
  mutate(sd = sqrt(mu * (1 - mu) / (1 + phi)))
```

### Mean

Conditional means (for "average" person on "average" scenario) on last trial:

```{r}
plot_means = eval_period_ends %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = mu, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 0.8) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid()
  #coord_cartesian(xlim = c(0.85, 1.0), ylim = c(1, 10.5)) 
plot_means %>%
  ggsave(filename = "img/rq/rq4-mu.png",
         height = 5,
         width = 6)

plot_means
```


### Mean differences

Mean differences by evaluation periods

```{r}
mean_diff = eval_period_ends %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"ep30","ep1")) %>%
  mutate(evalPeriod = factor(evalPeriod, levels = c("ep30","ep1"))) %>%
  mutate(evalPeriod = fct_relevel(evalPeriod, "ep1")) %>%
  select(treatment, mu, evalPeriod, .iteration) %>%
  pivot_wider(
    names_from = evalPeriod,
    values_from = mu
  ) %>%
  mutate(mu_diff = ep30 - ep1) %>%
  ggplot(aes(y = treatment, x = mu_diff)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 0.8) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  cowplot::theme_minimal_vgrid() +
  xlim(-.25,.25)
  #coord_cartesian(xlim = c(0.85, 1.0), ylim = c(1, 10.5)) 

mean_diff
```


### Standard deviation

Conditional standard deviation (for "average" person on "average" scenario) on last trial:

```{r}
plot_sd = eval_period_ends %>%
  mutate(evalPeriod = if_else(eval_period_normalized==0.5,"30","1")) %>%
    mutate(evalPeriod = factor(evalPeriod, levels = c("30","1"))) %>%
  ggplot(aes(y = treatment, x = sd, group = evalPeriod, fill = evalPeriod)) +
  geom_halfeyeh(fun.data = median_qih, fatten.point = 0.8) +
    scale_fill_OkabeIto(alpha = 0.5) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  cowplot::theme_minimal_vgrid() 
  #coord_cartesian(xlim = c(0.85, 1.0), ylim = c(1, 10.5)) 
plot_sd %>%
  ggsave(filename = "img/rq/rq4-std.png",
         height = 5,
         width = 6)

plot_sd
```


```{r, fig.width = 8, fig.height = 4, include = FALSE}
library(pBrackets) 

bracketsGrob <- function(...){
l <- list(...)
e <- new.env()
e$l <- l
  grid:::recordGrob(  {
    do.call(grid.brackets, l)
  }, e)
}

b1 <- bracketsGrob(0.5, 0.05, 0, 0.05, h=0.03, lwd=1, col="black")
b2 <- bracketsGrob(0.5, 0.05, 1, 0.05, h=0.13, lwd=1, col="black")

rates <- round(seq(0.6,1,0.1)*optimal_expreturn,4)*100

labs_vis = c("Point + Interval","Bar Chart","Interval","Table","HOP + Strip","Dot plot","HOP","Probability Density")
names(labs_vis) = vis_display_order

stock_labels <- c(20,40,60,80,100)
#dv_returns[round(dv_returns$dv==0.6,]

last_trial_plots = plot_grid(
  plot_means +
    ylab(NULL) +
    scale_y_discrete(labels = labs_vis) +
    annotate("text", label = "Myopic loss aversion", x = 0.82, y = 8.75, size = 2.5, color = "black") +
    annotate("text", label = "Equity premium", x = 0.88, y = 1.45, size = 2.5, color= "black") +
    annotate("text", label = "Optimal", x = 0.938, y = .6, size = 2.5, color="black") +
    annotation_custom(b1, xmin = 0.75, xmax = 1.25, ymin = 1.5) +
    annotation_custom(b2, xmin = 0.73, xmax = 0.92, ymin = 8.4) +
      annotate("label", label = "1 year", x = 0.72, y = 7.55, size = 2.5, color = "#56B4E9", alpha = 0.8, fontface =2, label.size = 0.5) +
        annotate("label", label = "30 years", x = 0.95, y = 5.4, size = 2.5, color = "#E69F00", alpha = 0.8, fontface =2, label.size = 0.5) +
    annotate("segment",  x = 0.973, y = 0.6, xend = 1, yend = 0.6, color = "black", arrow=arrow(length = unit(0.05,"inches"))) +
    #geom_bracket(xmin = "0.925", xmax = "2.0", y.position = 1.3, label = "Distance B") +
    xlab("") +
     scale_x_continuous(breaks = seq(0.6,1,0.1), labels = paste0(stock_labels,"%\n",rates,"%\n",round(seq(0.6,1,0.1),1),"\n",getExpectedInvestment(seq(0.6,1,0.1),50000,optimal_expreturn))) +
    theme(
      #plot.tag.position = c(0.05, 0.02),
      plot.background = element_rect(fill = 'white', color = 'white'),
      axis.text.x = element_text(size = 7),
      axis.line.y = element_blank(), 
      axis.ticks.y = element_blank(),
      legend.position = "none"),
  mean_diff +
    scale_y_discrete(labels = NULL) +
    ylab(NULL) +
    xlab("Difference by evaluation period\nMyopic loss aversion") +
    theme(
      axis.line.y = element_blank(), 
      axis.text.x = element_text(size = 8),
      axis.title.x = element_text(size = 8),
      axis.ticks.y = element_blank(),
      plot.background = element_rect(fill = 'white', color = 'white'),
      #legend.key.size = 8,
      legend.key.size = unit(0.2, 'cm'),
      legend.title=element_text(size=8),
      legend.text=element_text(size=8),
      legend.position = "none")
  ,
  plot_sd +
    scale_y_discrete(labels = NULL) +
    ylab(NULL) +
    xlab("Standard deviation") +
    annotate("label", label = "30 years", x = 0.05, y = 5.5, size = 2.5, color = "#E69F00", alpha = 0.8, fontface =2, label.size = 0.5) +
    annotate("label", label = "1 year", x = 0.155, y = 7.46, size = 2.5, color = "#56B4E9", alpha = 0.8, fontface =2, label.size = 0.5) +
    labs(fill = "Evaluation\nperiod") +
    #scale_fill_discrete(name = "Evaluation\nperiod") +
    theme(
      axis.line.y = element_blank(), 
      axis.text.x = element_text(size = 8),
      axis.ticks.y = element_blank(),
            axis.title.x = element_text(size = 8),
      #legend.key.size = 8,
      legend.key.size = unit(0.2, 'cm'),
      legend.title=element_text(size=8),
      legend.text=element_text(size=8),
      legend.position = "none") +
    theme(
            plot.background = element_rect(fill = 'white', color = 'white')
    ), # change to put in key
  # plot_sd_vs_text + 
  #   scale_y_discrete(labels = NULL) +
  #   ylab(NULL) +
  #   xlab("sd - text sd, last trial") +
  #   theme(axis.line.y = element_blank(), axis.ticks.y = element_blank()),
  # plot_sd_vs_dot50 + 
  #   scale_y_discrete(labels = NULL) +
  #   ylab(NULL) +
  #   xlab("sd - dot50 sd, last trial") +
  #   theme(axis.line.y = element_blank(), axis.ticks.y = element_blank()),
  align = "h",
  ncol = 4,
  rel_widths = c(0.1,2.5,1.25,1.25),
  axis = "b"
)

small_label ="Stock allocation decision\nExpected rate of return\nExp. return / optimal return (DV)\nHypothetical value at retirement"

last_trial_plots = 
  cowplot::ggdraw(last_trial_plots) + draw_label(small_label, x = 0.09, y = 0.136, size = 7)

last_trial_plots
pdf(file = "img/rq/eval_end_plots.pdf", useDingbats = FALSE, width = 8, height = 4)
last_trial_plots
dev.off()
```

